<<<<<<< HEAD
Peer Assessment 1
========================================================

Loading and preprocessing the data

```{r}
echo = TRUE
library(ggplot2)
setwd("/Users/jamessweeney2/Desktop/Coursera")

## Load the data
activity <- read.csv("activity.csv")

## Remove NA data
act <- activity[!is.na(activity$steps),]
```

What is the Mean total number of steps taken per day?

```{r}
## Create hisogram of steps per day
g <- ggplot(data = act, aes(x = steps))
g + geom_histogram(binwidth = 10) + coord_cartesian(ylim = c(0, 500))

## Calculate mean and median steps per day and print
means <- tapply(act$steps, act$date, mean)
medians <- tapply(act$steps, act$date, median)
print(means, digits = 0)
print(medians, digits = 0)
```

What is the average daily activity pattern?

```{r}
## find means by time
hours <- tapply(act$steps, act$interval, mean)
hours <- cbind(hours, names(hours))
colnames(hours) <- c("steps", "interval")

## Plot means by time
plot(hours[,2], hours[,1], type = "l", xlab = "Time", ylab = "Steps", main = "Average Steps by Time of Day")

## Calculate which time had most steps on average
hours[which.max(hours[,1]),2]
```

Imputing missing values

```{r}

## Count missing values
sum(!complete.cases(activity))

## Create a function to count the data without NAs
R <- function(x) {sum(complete.cases(x))}

## Apply that function over activity data
data <- tapply(activity$steps, activity$date, R)

## Show data available data by date
check <- cbind(data, means)

## List dates with no data
zeros <- row.names(check[check[,1] == 0,])

## show averages by interval in numeric format
intaverages <- cbind(as.numeric(hours[,1]), activity[1:288,3])

## fill in NAs with average daily data in activity2
activity2 <- activity
for(i in 1:length(zeros)) {
    activity2[activity2[,2] == zeros[i],][,1] <- intaverages[,1]
}

## Plot histogram of filled in data
h <- ggplot(data = activity2, aes(x = steps))
h + geom_histogram(binwidth = 10) + coord_cartesian(ylim = c(0, 1000))

# Calculate means and medians of filled in data
means2 <- tapply(activity2$steps, activity2$date, mean)
medians2 <- tapply(activity2$steps, activity2$date, median)
print(means2, digits = 0)
print(medians2, digits = 0)
```

Are there differences in activity patterns between weekdays and weekends?

```{r}
## Create a factor variable in activity2 for whether it is a weekend
day <- cbind(weekdays(as.Date(activity[,2])), "weekday")
day[day[,1] == "Saturday",][,2] <- "weekend"
day[day[,1] == "Sunday",][,2] <- "weekend"
activity2[,4] <- as.factor(day[,2])
names(activity2) <- c("steps", "date", "interval", "weekend")

bytime <- tapply(activity2$steps, activity2$interval, mean)
bytime <- cbind(bytime, names(bytime))
colnames(bytime) <- c("steps", "interval")

## split activity2 data into weekend and weekday data
tt <- split(activity2, activity2$weekend)
weekendtime <- tapply(tt$weekend[,1], tt$weekend[,3], mean)
weekdaytime <- tapply(tt$weekday[,1], tt$weekday[,3], mean)
weekendtime <- cbind(weekendtime, names(weekendtime))
weekdaytime <- cbind(weekdaytime, names(weekdaytime))

## Plot steps by interval for weekdays and weekends
par(mfrow = c(2,1))
plot(as.numeric(as.character(weekendtime[,2])), as.numeric(as.character(weekendtime[,1])), type = "l", ylab = "Steps", xlab = "Interval", main = "Weekend Steps by Interval")
plot(as.numeric(as.character(weekdaytime[,2])), as.numeric(as.character(weekdaytime[,1])), type = "l", ylab = "Steps", xlab = "Interval", main = "Weekday Steps by Interval")
```



=======
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> upstream/master
